---
title: "CAPSTONE-PH125.9x"
author: "Laxmansingh Rajput"
date: "5/31/2019"
output:
  pdf_document: default
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE, 
                      message = FALSE, 
                      cache = TRUE, 
                      comment = NA)

library(tidyverse)
library(kableExtra)
library(knitr)
library(ggplot2)

```
## OBJECTIVES

Creating a movie recommendation system using the MovieLens dataset. Train a machine learning algorithm using the inputs in one subset to predict movie ratings in the validation set.

This project is being completed in R version 3.6.0 (_Planting of a Tree_)

The submission for the MovieLens project will be three files: 

*  A report in the form of an Rmd file
*  A report in the form of a PDF document knit from your Rmd file 
*  An Rmd/R script file that generates the predicted movie ratings and calculates RMSE

The movie rating predictions will be compared to the true ratings in the validation set using RMSE and the report outputs the RMSE.

## EXECUTIVE SUMMARY  
    1. Dataset
       a. MovieLens dataset (downloaded as R Version is 3.6.0)
          from the following google drive 
          <https://drive.google.com/drive/folders/1IZcBBX0OmL9wu9AdzMBFUG8GoPbGQ38D>
    2. Goals of the project
       a. Predict the ratings for the validation dataset
       b. Compute the RMSE
    3. Key Steps
       a. Compute the average of the movie ratings
       b. Calculate the reqularized parameter (lambda) using cross-validation
       c. Get the lambda that gives the least RMSE
       d. Use the lambda to predict the ratings
       e. Compute the RMSE on the predict ratings vs the validation test

\newpage

```{r Read edx set and Validation set, echo=FALSE, message=FALSE}

###################################
# Read edx set and validation set #
###################################

set.seed(1, 
         sample.kind = "Rounding") # if using R 3.6.0

edx <- readRDS("Data/edx.rds")
validation <- readRDS("Data/validation.rds")

tab_rating <- data.frame(table(edx$rating))
names(tab_rating) <- c("Rating", "Frequency")

##
##  Top 5 rated movies
##

top_5rated_movies <- edx %>% 
                        group_by(title) %>% 
                        tally() %>% 
                        arrange(., desc(n)) %>% 
                        head(.,5)

names(top_5rated_movies) <- c("Title", "Frequency")


ratings_in_desc <- edx %>% 
                      group_by(rating) %>% 
                      tally() %>% 
                      arrange(., desc(n))

names(ratings_in_desc) <- c("Ratings", "Frequency")

DESCRIPTION <- c("Observations", 
                 "Variables", 
                 "Unique User IDs", 
                 "Unique Movie IDs", 
                 "Unique Ratings", 
                 "Mean Rating")

VALUES <- rbind(formatC(as.numeric(dim(edx)[1]),format="d"),
                    formatC(as.numeric(dim(edx)[2]),format="d"),
                    formatC(as.numeric(n_distinct(edx$userId)),format="d"),
                    formatC(as.numeric(n_distinct(edx$movieId)),format="d"),
                    formatC(as.numeric(n_distinct(edx$rating)),format="d"),
                    formatC(as.numeric(mean(edx$rating)),format="f",digits=6))
##
##  Create table with the dataset information
##
dataset_information <- data.frame(DESCRIPTION, 
                                  VALUES)


```

## METHOD/ANALYSIS SECTION
* Explore the data
* Explore the rating information
* Explore the top rated movies
* Model selection - use the regularization method on the rating based on movieID and userID biases

#### FUNCTIONS TO BE USED
* RMSE <- function(true_ratings, predicted_ratings)
               sqrt(mean((true_ratings - predicted_ratings)^2))

* predict_rating <- compute the movie and user biases, and return the predicted ratings based on the
                     regularization parameter.


```{r Functions, echo=FALSE, message=FALSE}

##
##  Root mean square error function
##
RMSE <- function(true_ratings, predicted_ratings){
        sqrt(mean((true_ratings - predicted_ratings)^2))
}
##
##  Predicted rating based on the movie and user biases
##
predict_rating <- function(reg_param){
  ##
  ##  movie bias based on movieID (group_by) 
  ##
  movie_bias <- edx %>% 
    group_by(movieId) %>%
    summarize(movie_bias = sum(rating - mu)/(n()+reg_param))  # movie bias 
  ##
  ##  user bias based on userID (group_by) 
  ##     movie bias also being taken into consideration
  ##
  user_bias <- edx %>% 
    left_join(movie_bias, 
              by="movieId") %>%
    group_by(userId) %>%
    summarize(user_bias = sum(rating - movie_bias - mu)/(n()+reg_param)) # User bias
  ##
  ##  return the prediction from the function
  ##
  return(validation %>% 
             left_join(movie_bias, 
                       by = "movieId") %>%
             left_join(user_bias, 
                       by = "userId") %>%
             mutate(pred = mu + movie_bias + user_bias) %>%     # predictions for the movies
             pull(pred))                                        # extract only the prediction 
  
}



```


#### EXPLORE THE DATASET

We can examine the edx data by using R's `glimpse` command:

```{r glimpse}
glimpse(edx)
```

Here we see that we have a data frame, accessible at `edx`:

```{r head(edx)}
head(edx, 6)
```

\newpage

#### Understanding Variables 

1. The list of the variables.

```{r Explore the dataset, echo=FALSE}

kable(t(variable.names(edx)), 
      caption = "Variable Names") %>%  
  kable_styling(bootstrap_options = c("striped", "hover"), 
                full_width = F, latex_options = "hold_position")

```

2. Other information regarding the dataset - Observations, variables, unique values, and mean rating

```{r Variable Information, echo=FALSE}
options(scipen=999)

kable(dataset_information, align = c("l", "r"),
      caption = "Variable Information") %>%  
  kable_styling(bootstrap_options = c("striped", "hover"), 
                full_width = F, 
                latex_options = "hold_position")

ggplot(tab_rating, aes(Rating, Frequency)) + 
  geom_bar(stat = "identity", fill = "blue")  +
  ggtitle("Rating Information")
  
```

\newpage

3. Rating information 

```{r Rating Information, echo=FALSE}

kable(t(tab_rating), align=rep('c', 7),
      caption = "Rating Information") %>%  
  kable_styling(bootstrap_options = c("striped", 
                                      "hover"), 
                full_width = F, 
                latex_options = "hold_position") %>%
  footnote(general = "   Overall Summary of the ratings. ",
           number = c("     Rating ranges from 0.5 to 5 in 0.5 steps "))
```

```{r Rating in reverse order, echo=FALSE}

kable(ratings_in_desc, 
    caption = "Rating in reverse order") %>%  
  kable_styling(bootstrap_options = c("striped", "hover"), 
                full_width = F, latex_options = "hold_position")

```


4. Five most rated movies

```{r Top 5 rated movies, echo=FALSE}

kable(top_5rated_movies, 
      caption = "Top 5 rated movies") %>%  
  kable_styling(bootstrap_options = c("striped", "hover"), 
                full_width = F, 
                latex_options = "hold_position")
```

\newpage

5. edx dataset

```{r edx dataset, echo=FALSE}
kable(select(head(edx, 20),1,2,3,5,6),
      caption = "edx Dataset (Snapshot)") %>%  
  kable_styling(bootstrap_options = c("striped", "hover"), 
                full_width = F, 
                latex_options = c("scale_down","hold_position"))

```

\newpage

6. validation dataset

```{r validation dataset, echo=FALSE}
kable(select(head(validation, 20),1,2,3,5,6),
      caption = "validation Dataset (Snapshot)") %>%  
  kable_styling(bootstrap_options = c("striped", "hover"), 
                full_width = F, 
                latex_options = c("scale_down","hold_position"))

```

\newpage

##   CROSS-VALIDATION TO PICK A REGULARIZATION PARAMETER (LAMBDA)
####             MOVIE + USER EFFECT
    Use sapply function to return a vector and then retrieve the 
      parameter that gives the minimum RMSE
         1. Calculate the regularized bias for each movies
         2. Calculate the regularized bias for each user
         3. Predict ratings based on both regularized movies & user biases
         4. Compare against the validation (test) set and compute the RMSE


```{r cross-validation to pick a regularization paramater (lambda), echo=FALSE}
##
##   use cross-validation to pick a λ:
##
lambdas <- seq(0, 15, 0.25)

mu <- mean(edx$rating)


rmses <- sapply(lambdas, function(reg_param){

  predicted_ratings <- predict_rating(reg_param)          # both of the functions could be combined,
    
  return(RMSE(predicted_ratings, validation$rating))      # but I kept it separate for readability
})

## paste(lambdas, rmses)

qplot(lambdas, 
      rmses)  
```

\newpage

##  RESULTS SECTION
*  Best parameter
*  Regularization parameters and RMSE

```{r RESULTS SECTION, echo=FALSE, fig.height=10, fig.width=7}

paste("Best regularization parameter: ", 
      lambdas[which.min(rmses)])

lambdas_rmses <- data.frame(lambdas, rmses)
names(lambdas_rmses) <- c("LAMBDAS", "RMSE")

library(gridExtra)
library(grid)

x <- lambdas_rmses[1:31,]
y <- lambdas_rmses[32:61,]
##
##   Using grid arrange so that we can display all the information
##
grid.arrange(tableGrob(x), 
             tableGrob(y), 
             ncol = 2,
             top = textGrob("Regularization parameters and RMSE",
                            x = 0, # starts far left
                            y = 0.5, # experiment with vertical placement
                            just = "left", # left-aligned,
                            gp = gpar(fontsize = 18) # bigger font
             ))


```

\newpage

##  CONCLUSION SECTION
*  Best RMSE
*  Some predicted ratings 

```{r CONCLUSION SECTION, echo=FALSE}
##
##   Select the best regularization parameter to get the prediction
##       and compute RMSE based on the prediction against the validation dataset
##
reg_param <- lambdas[which.min(rmses)]  
predicted_ratings <- predict_rating(reg_param)
    
paste("Best RMSE: ", RMSE(predicted_ratings, validation$rating))

```

## PREDICTED RATING (SNAPSHOT)

```{r Predicted Ratings, echo=FALSE}
kable(select((cbind(validation, predicted_ratings) %>% head(20)), 1,2,3,5,7),
      caption = "PREDICTED RATINGS (Snapshot)") %>%  
  kable_styling(bootstrap_options = c("striped", "hover"), 
                full_width = F, latex_options = "hold_position")

```

\newpage

##  Information About The Current R Session

```{r SESSION INFORMATION SECTION, echo=FALSE}

sessionInfo()


```